resevol: an R package for spatially explicit models of pesticide resistance

### 2023-07-17 notebook/presentation/resevol_presentation.Rmd
Presentation for British Ecological Society 2022

---
title: "resevol: an R package for spatially explicit models of pesticide resistance
  given evolving pest genomes"
author: |
  | A. Bradley Duthie$^{1}$, Rosie Mangan$^{1}$, C. Rose McKeon$^{1}$,
  | Matthew C. Tinsley$^{1}$, and Luc F. Bussi&egrave;re$^{2,3}$
date: '[1] University of Stirling [2] University of Gothenburg [3] Gothenburgh Global Biodiversity Centre alexander.duthie@stir.ac.uk, @bradduthie@ecoevo.social http://bradduthie.github.io/talks/BES2022.pdf'
output:
  beamer_presentation:
    theme: "default"
    colortheme: "default"
    fonttheme: "default"
  ioslides_presentation: default
  slidy_presentation: default
header-includes:
- \usepackage{hyperref}
- \usepackage{amsmath}
colorlinks: true
linkcolor: blue
urlcolor: blue
---

<!---

ioslides should save these as a pdf. To make it easy for students, use pdfjam

$ sudo apt install texlive-extra-utils
$ pdfjam --batch --nup 2x3 --suffix 1up --outfile . filename.pdf

NOTE: Need to alternate slides with lines for note-taking

--->

@@

\vspace{2mm}

@@

\begin{itemize}
\item Vary biopesticide application$^{3,4}$
\item Polygenic resistance traits$^{5}$
\item Negative cross-resistance$^{6}$
\end{itemize}

@@

\footnotetext[1]{Gould et al. 2018. \textit{Ecol. Appl.} 13:1791-1805.}
\footnotetext[2]{Bradshaw et al. 2016. \textit{Nat. Commun.} 7:12986.}
\footnotetext[3]{Carroll et al. 2014. \textit{Science} 346:1245993.}
\footnotetext[4]{Saikai et al. 2021. \textit{Pest Manag. Sci.} 77:273-284.}
\footnotetext[5]{Boots, M. 2011. \textit{Am. Nat.} 178:214-220.}
\footnotetext[6]{Via, S. 1986. \textit{Pesticide resistance}. 222-235.}

## Agricultural heterogeneity to overcome resistance evolution

@@

\begin{columns}
\begin{column}{0.5\textwidth} 
\begin{center}
Unique farms
\end{center}
\end{column}
\begin{column}{0.5\textwidth} 
\begin{center}
Biopesticide applied
\end{center}
\end{column}
\end{columns}

@@

@@

@@

\begin{columns}
\begin{column}{0.5\textwidth} 
\textbf{Complex individuals}
\begin{itemize}
\item Complete genomes
\item Highly polygenic traits
\item Set trait covariances
\item Flexible life-history
\item Flexible mating system
\end{itemize}
\end{column}
\begin{column}{0.5\textwidth} 
\textbf{Dynamic landscapes}
\begin{itemize}
\item Custom land dimensions
\item Custom pesticide rotation
\item Custom crop rotation
\item Individual farms tracked
\item Raster landscape maps
\end{itemize}
\end{column}

@@

@@

\vspace{3mm}

@@

\begin{itemize}
\item Simulations of pest ecology, resistance evolution, and agricultural production
\item Covarying polygenetic traits
\item Flexible pest life-histories, mating systems, traits
\item Highly customisable, detailed landscapes
\item Free to use$^{1}$, open source$^{2}$ (GPL >= 2), documented$^{1,3}$, actively maintained$^{4}$
\end{itemize}

\footnotetext[1]{\textbf{Website}: \href{https://bradduthie.github.io/resevol/}{https://bradduthie.github.io/resevol/}}
\footnotetext[2]{\textbf{Code}: \href{https://github.com/bradduthie/resevol}{https://github.com/bradduthie/resevol}}
\footnotetext[3]{\textbf{Lab Notebook}: \href{https://bradduthie.github.io/resevol/notebook/}{https://bradduthie.github.io/resevol/notebook/}}
\footnotetext[4]{\textbf{Email}: alexander.duthie@stir.ac.uk \textbf{Mastodon}: @bradduthie@ecoevo.social}

>>>>

---
title: "resevol: an R package for spatially explicit models of pesticide resistance
  given evolving pest genomes"
author: |
  | A. Bradley Duthie$^{1}$, Rosie Mangan$^{1}$, C. Rose McKeon$^{1}$,
  | Matthew C. Tinsley$^{1}$, and Luc F. Bussi&egrave;re$^{2,3}$
date: '[1] University of Stirling [2] University of Gothenburg [3] Gothenburgh Global Biodiversity Centre alexander.duthie@stir.ac.uk, @bradduthie@ecoevo.social http://bradduthie.github.io/talks/resevol.pdf'
output:
  beamer_presentation:
    theme: "default"
    colortheme: "default"
    fonttheme: "default"
  ioslides_presentation: default
  slidy_presentation: default
header-includes:
- \usepackage{hyperref}
- \usepackage{amsmath}
colorlinks: true
linkcolor: blue
urlcolor: blue
---

@@

\vspace{8mm}\pause

@@

\begin{itemize}
\item Weaken selection for resistance$^{3}$
\item Varying pesticide application$^{4}$
\end{itemize}

@@

\footnotetext[1]{Gould et al. 2018. \textit{Ecol. Appl.} 13:1791-1805.}
\footnotetext[2]{Bradshaw et al. 2016. \textit{Nat. Commun.} 7:12986.}
\footnotetext[3]{Carroll et al. 2014. \textit{Science} 346:1245993.}
\footnotetext[4]{Saikai et al. 2021. \textit{Pest Manag. Sci.} 77:273-284.}

## Negatively correlated cross-resistance between pesticides

```{r, echo = FALSE, fig.width = 4.65, fig.height = 4.65}
var1 <- runif(n = 120, min = 0, max = 1);
var2 <- rnorm(n = 120, sd = 0.2) - var1;
par(mgp=c(0.2,1,0))  
plot(x = var1, y = var2, xaxt = "n", yaxt = "n", pch = 20, cex = 1.25,
     xlab = "Resistance to pesticide A", ylab = "Resistance to pesticide B",
     cex.lab = 1.25);
par(mgp=c(3,1,0))
```

\footnotetext[1]{Chapman \& Penman. 1979. \textit{Nature} 281:298-299.}

## Enhancing Diversity to Overcome Resistance Evolution

@@

\begin{columns}
\begin{column}{0.5\textwidth} 
\begin{center}
Unique farms
\end{center}
\end{column}
\begin{column}{0.5\textwidth} 
\begin{center}
Pesticide applied
\end{center}
\end{column}
\end{columns}

@@

```{r, out.width = "100%", echo = FALSE}
include_graphics("img/Endorse_people.png");
```

@@

\begin{columns}
\begin{column}{0.5\textwidth} 
\textbf{Complex individuals}
\begin{itemize}
\item Complete genomes
\item Highly polygenic traits
\item Set trait covariances
\item Flexible life-history
\item Flexible mating system
\end{itemize}
\end{column}
\pause
\begin{column}{0.5\textwidth} 
\textbf{Dynamic landscapes}
\begin{itemize}
\item Custom land dimensions
\item Custom pesticide rotation
\item Custom crop rotation
\item Individual farms tracked
\item Raster landscape maps
\end{itemize}
\end{column}

@@

@@

\begin{columns}
\begin{column}{0.5\textwidth} 
\textbf{Pre-specified covariances}

\begin{align}
\begin{bmatrix}
 1.0 & -0.5 \\
-0.5 &  2.0
\end{bmatrix}
\end{align}

\end{column}
\begin{column}{0.5\textwidth} 
\textbf{Covariances produced}

\begin{align}
\begin{bmatrix}
 1.007 & -0.436 \\
-0.436 &  1.906
\end{bmatrix}
\end{align}

@@

\vspace{3mm}\pause

@@

\begin{itemize}
\item Rapidly develop and simulation models of pest ecology and resistance evolution
\item Introduces a novel approach to modelling complex pest genetic architecture
\item Flexible pest life-histories, mating systems, trait covariances
\item Highly customisable, detailed landscapes
\item Free to use$^{1}$, open source$^{2}$ (GPL >= 2), documented$^{1,3}$
\end{itemize}

\footnotetext[1]{Website: \href{https://bradduthie.github.io/resevol/}{https://bradduthie.github.io/resevol/}}
\footnotetext[2]{Code: \href{https://github.com/bradduthie/resevol}{https://github.com/bradduthie/resevol}}
\footnotetext[3]{Lab Notebook: \href{https://bradduthie.github.io/resevol/notebook/}{https://bradduthie.github.io/resevol/notebook/}}
